---
title: "Regression "
author: "Martin Sigrist (21-101-175)"
date: "03.06.2023"
output: html_document
---

# Introduction

The aim of this chapter is stepwise forward regression. The dependent variable is the gross primary production (GPP). It is defined a the gross CO~2~ uptake flux driven by photosynthesis.

The used data is from the Fluxnet from the Davos site (https://fluxnet.org/sites/siteinfo/CH-Dav). The following variables are considered in for the evaluation:

-   TA\_\* (°C): Air temperature
-   SW_IN\_\* (W m^−2^): Shortwave incoming radiation
-   LW_IN\_\* (W m^−2^): Longwave incoming radiation
-   VPD\_\* (hPa): Vapor pressure deficit (the difference between actual and saturation water
    vapor pressure)
-   PA\_\* (kPa): Atmospheric pressure
-   P\_\* (mm): Precipitation
-   WS\_\* (m s^−1^): Wind speed
-   CO2\_\* (µmolCO~2~ mol^−1^): Carbon Dioxide (CO~2~) mole fraction in moist air
-   PPFD_IN (µmolPhotons m^−2^ s^−1^): incoming photosynthetic photon flux density
-   GPP\_\* (μmol~CO2~ m^−2^ s^−1^): Gross primary production (the ecosystem-level gross CO2 uptake flux driven by
    photosynthesis)

# Load required libraries

```{r, message=FALSE}
library(tidyverse)
```

# Load data

```{r}
half_hourly_fluxes <- read_csv("../data/df_for_stepwise_regression.csv", show_col_types = FALSE)
```

# Bivariate Models

In a first step, a bivariate model for each independent variable is calculated

```{r, message=FALSE}
# Vector with names of independent variables
indep_vars <- colnames(half_hourly_fluxes)[! colnames(half_hourly_fluxes) %in% c("siteid", "TIMESTAMP", "GPP_NT_VUT_REF")]

# Create empty variable for results
res <- NULL 

for (i in indep_vars) {
  # Calculate linear model
  linmod <- lm(as.formula(paste("GPP_NT_VUT_REF ~ ", i)), data = half_hourly_fluxes)
  
  # Add results for actual model to result data frame
  res <- bind_rows(res,
                   tibble(
                     feature = i,
                     mse = mean(linmod$residuals^2),
                     R2 = summary(linmod)$r.squared,
                     R2_adj = summary(linmod)$adj.r.s,
                     AIC = extractAIC(linmod)[2],
                     BIC = BIC(linmod))) # this implementation is based on log-likelihood
}

# Show results, rounded on three decimal places
res |> mutate_at(vars(-feature), ~round(., 3)) |>
  knitr::kable(format = "html") |>
  kableExtra::kable_classic(full_width = F)
```

## Scatter plots

```{r}
# Create scatter plots for each independent variable
half_hourly_fluxes |>
  select_if(is.numeric) |>
  gather(-GPP_NT_VUT_REF, key = "var", value = "value") |>
  drop_na() |>
  ggplot(aes(x = value, y = GPP_NT_VUT_REF)) +
    geom_point(alpha = 0.3) +
    stat_smooth(formula = y ~ x, method = "lm", color = "red", se = FALSE) +
    facet_wrap(~ var, scales = "free") +
    theme_bw()
```

## Discussion

R^2^ reaches the highest values in the bimodal linear regressions for the air temperature (TA_F, TA_F_MDS), the incoming short wave radiation (SW_IN_F, SW_IN_F_MDS) and the incoming photosynthetic photon flux density (PPFD_IN). Variables with the same name but with different suffixes (_F, _F_MDS) are principally the same measurement, but for the first, gaps are filled by Marginal Distribution Sampling (MDS) (Pastorello et al. 2020). The resulting values for each variable pair are therefore very similar.
The determination coefficient and the mean square error for the PPFD_IN variable are similar to them of the short wave radiation. The photosynthetically active radiation is part of the total incoming radiation. Similar results are expected and observed.

# Stepwise forward regression

In the second part, a linear model is compiled stepwise. For each step, one more independent variable is added to the model until the Akaike’s Information Criterion (AIC) reaches a minimal value.

```{r, message=FALSE}
# Vector with names of independent variables
indep_vars <- colnames(half_hourly_fluxes)[! colnames(half_hourly_fluxes) %in% c("siteid", "TIMESTAMP", "GPP_NT_VUT_REF")]

# Create empty data frame for results
df_res <- data.frame(matrix(nrow = 0, ncol = length(indep_vars) + 2))
#provide column names
colnames(df_res) <- append(append(c("base.model"), indep_vars),"AIC")

model_vars <- NULL # Define empty variable for the independent variables of the actual model
step_AIC <- NULL # Variable to store AIC value for the highest R2 in the sctual step

n_step <- 0 # Variable for if condition in the repeat-loop to avoid index problems

repeat { # Loop to go through the independent variables
  max_R2 <- 0 # Variable to store the highest R2 for the actual step
  
  # Add new row to result data frame
  df_res[nrow(df_res)+1,] <- rep(NA, 16) # Fill new row with N/As
  df_res[nrow(df_res),1] <- paste(model_vars, collapse=" + ") # Fill first cell in new row with the previous model formulation
  

  for (i in indep_vars) { # Loop to go though the remaining variables, not yet in the model
    # Calculate linear model (Remove all N/As to become comperable with the result from the step function, that only works without N/As)
    linmod <- lm(as.formula(paste("GPP_NT_VUT_REF ~ ", paste(append(model_vars, i), collapse=" + "))), data = half_hourly_fluxes |> drop_na())
    
    # Store actual predictor, AIC and R2 if R2 is the highest in the actual round
    if(summary(linmod)$r.squared > max_R2) {
      max_R2 <- summary(linmod)$r.squared
      step_AIC <- extractAIC(linmod)[2]
      best_predictor <- i
    }
    
    # Add R2 for the actual variable to the result data frame
    df_res[nrow(df_res),i] <- summary(linmod)$r.squared
  }
  
  # Add best predictor for the actual step to model
  model_vars <- append(model_vars, best_predictor)
  
  # Add AIC for the highest R2 in the actual step to the result data frame
  df_res[nrow(df_res),"AIC"] <- step_AIC
  
  # Remove best predictor from actual round from overall list of independent variables
  indep_vars <- indep_vars[!indep_vars %in% best_predictor]

  
  if(step_AIC > df_res[nrow(df_res)-n_step,"AIC"]) { # Break condition for repeat loop
    break
  }
  n_step <- 1
}

# Print best model
cat("Best model: ", paste("GPP_NT_VUT_REF ~ ", df_res[nrow(df_res),1]), "\n")
cat("R2 for best model: ", max(df_res[nrow(df_res)-1,3:ncol(df_res)-1], na.rm = TRUE), "\n")
cat("AIC for best model: ", df_res[nrow(df_res)-1,ncol(df_res)], "\n")
```

## Results Visualisation

```{r}
plot(apply(df_res[,3:ncol(df_res)-1], 1, max, na.rm = TRUE),
     xlab = "Step",
     ylab = bquote(italic(R)^2),) # Max value of Coefficient of determination for each step
```

Plotting R^2^ against each step of the stepwise forward regression shows its convergence.

# Using step-funtcion from stats-library

```{r}
# Generate scope model for the stepwise forward regression function "step"
scope_model <- paste("GPP_NT_VUT_REF ~ ", paste(colnames(half_hourly_fluxes)[! colnames(half_hourly_fluxes) %in% c("siteid", "TIMESTAMP", "GPP_NT_VUT_REF")], collapse=" + "))

stats::step(
  object = lm(GPP_NT_VUT_REF ~ 1, # Start with the simplest model with only a consant
                data = half_hourly_fluxes |> select(-siteid, -TIMESTAMP)|> drop_na() # Remove all NA to make the step function work
              ),
  scope = scope_model, # Scope model, including all variables
  direction = "forward"
)
```

# Discussion

The stepwise forward regression algorithm is working. The implementation would accept also a different number of independent variables. Future improvements for an increased efficiency might be possible. The computational effort was not a scope of this work, as the amount of data to be processed is kept rather low. The step-function from the stats-library gives the same results for the final linear model, R^2^ and AIC.

In the final model, from the two temperature variables (TA_F and TA_F_MDS) and the two incoming short wave radiation (SW_IN_F and SW_IN_F_MDS) are only one present. Because two of each represent the same measurements, they are not linear independent and adding the second one would not improve the model. A similar effect is observed between the PPFD_IN and the SW_IN_F variables. In the bivariate models, they both had a similar determination coefficient. But the wave length spectrum covered by the PPFD_IN variable is part of the wave length spectrum of the SW_IN_F variable and they are not fully independent. Therefore, the addition of the second one to the model has only a small positive effect on the determination coefficient.


# Literature

Pastorello, Gilberto, et al. 2020: The FLUXNET2015 dataset and the ONEFlux processing pipeline for eddy covariance data. Scientific Data 7 (1): 225. DOI: https://doi.org/10.1038/s41597-020-0534-3
