---
title: "re_tidy"
author: "Martin Sigrist"
date: "2023-03-06"
output: html_document
---

# Data wrangling

## Manual data cleaning

-   Removed header text
-   Modified column headers
  -   time in years
  -   CO2 content in g C / m^2
-   Removed empty rows
-   Filled up empty cells in experiment column

# Load Packages

```{r}
library(tidyverse)
# library(tidyr)
# library(dplyr)
# library(lubridate)
# library(ggplot2)
```


# Import Data
```{r}
elev_CO2_exp <- readr::read_delim("../data/1249534s1-s6_cleaned.csv", delim = ";", show_col_types = FALSE) # Read data set from file
elev_CO2_exp
```
# Log-responce ration per experiment
Data are aggregated per experiment by calculating the mean value for the ambient (low) and elevated (high) CO2 concentration. As the development over time is not from interest, the data evaluation is done like that.

```{r}
experiment_data <- elev_CO2_exp |>                                          # Elevated CO2 Experiments data set
  select(experiment, ends_with("CO2_mean")) |>                              # Take columns from interest
  group_by(experiment) |>                                                   # Group data by experiment
  summarise(amb_CO2 = mean(amb_CO2_mean), inc_CO2 = mean(inc_CO2_mean)) |>  # Calculate mean values per experiment
  mutate(LRR = log(inc_CO2 / amb_CO2))                                      # Calculate Log-Response Ration
experiment_data
```
# Log-responce ration per time phase
The log-responce ration per phase in the experiment is calculated first for each data point. The number of data point for the three stages (early, mid and late) are different. The mean values of the log-responce ratio for each phase is calculated at the end to avoid over weighting of the data points in the phases with a low number of measurements. The temporal development during the experiments are not from interest.

```{r}
# Adding new column with the phase definition (early, mid, late)
elev_CO2_exp <-
  elev_CO2_exp |>                                    # Elevated CO2 Experiments data set
    filter(exp_time < 3) |>                          # Take rows where experiment runs less than 3 years
    mutate(exp_phase = "early") |>                   # Add column with experiment phase
    bind_rows(                                       # Combine with rows with experiments in mid-phase
      elev_CO2_exp |>                                # Elevated CO2 Experiments data set
        filter(exp_time >= 3 & exp_time <= 6) |>     # Take rows where experiment runs between 3 and 6 years
        mutate(exp_phase = "mid")                    # Add column with experiment phase
    ) |>
    bind_rows(                                       # Combine with rows with experiments in late
      elev_CO2_exp |>                                # Elevated CO2 Experiments data set
        filter(exp_time > 6) |>                      # Take rows where experiment runs longer than 6 years
        mutate(exp_phase = "late")                   # Add column with experiment phase
    )

# Calculating the log-response ration per time phase
phase_LRR <- elev_CO2_exp |>                         # Elevated CO2 Experiments data set
  select(exp_phase, ends_with("CO2_mean")) |>        # Take columns from interest
  mutate(LRR = log(inc_CO2_mean / amb_CO2_mean)) |>  # Calculate Log-Response Ration for each measurement
  group_by(exp_phase) |>                             # Group data by experiment
  summarise(LRR_mean = mean(LRR))                    # Calculate mean values per experiment

# Print table with results
knitr::kable(phase_LRR, digits = c(NA, 4))
```

