---
title: "Data Wrangling"
author: "Martin Sigrist (21-101-175)"
date: "03.06.2023"
output: html_document
---

# Manual data cleaning

The following manual modifications have been done with Excel:

-   Removed header text
-   Modified column headers
    -   time in years
    -   CO~2~ content in g~C~ m^-2^
-   Removed empty rows
-   Filled up empty cells in experiment column

The cleaned data was saved in a csv-file for usage with R.

# Load Packages

```{r, message=FALSE}
library(tidyverse)
```

# Import Data

```{r}
# Read data set from file
elev_CO2_exp <- readr::read_delim("../data/1249534s1-s6_cleaned.csv", delim = ";", show_col_types = FALSE)
```

# Log-response ration per experiment

Observations are aggregated per experiment by calculating the mean value for the ambient (low) and the elevated (high) CO~2~ concentration. The evaluation can be done like that, because the development over time is not from interest.

```{r}
# Calculate log-response ratio per experiment
experiment_data <- elev_CO2_exp |>                                          # Elevated CO2 Experiments data set
  select(experiment, ends_with("CO2_mean")) |>                              # Take columns from interest
  group_by(experiment) |>                                                   # Group data by experiment
  summarise(amb_CO2 = mean(amb_CO2_mean), inc_CO2 = mean(inc_CO2_mean)) |>  # Calculate mean values per experiment
  mutate(LRR = log(inc_CO2 / amb_CO2))                                      # Calculate Log-Response Ration

# Table with calculated results
experiment_data |>
  knitr::kable(digits = c(NA, 1, 1, 4),
               col.names = c("Experiment",
                             "Soil organic carbon @ ambient CO~2~ level (g~C~ m^-2^)",
                             "Soil organic carbon @ elevated CO~2~ level (g~C~ m^-2^)",
                             "Log-responce ratio"),
               format = "html") |>
  kableExtra::kable_classic(full_width = F) |>
  kableExtra::column_spec(2, width = "2in") |>
  kableExtra::column_spec(3, width = "2in")
```

# Log-response ration per time phase

The log-response ration per phase in the experiment is calculated first for each data point. The number of data points for the three stages (early, mid and late) are different. The mean values of the log-response ratio for each phase is calculated at the end to avoid over weighting of the data points in the phases with a low number of measurements. The temporal developments during the experiments are not from interest.

```{r}
# Adding new column with the phase definition (early, mid, late)
elev_CO2_exp <-
  elev_CO2_exp |>                                                    # Elevated CO2 Experiments data set
    filter(exp_time < 3) |>                                          # Take rows where experiment runs less than 3 years
    mutate(exp_phase_name = "early phase (<3 years since start)",    # Add column with experiment phase name
           exp_phase_num = 1) |>                                     # Add column with experiment phase number
    bind_rows(                                                       # Combine with rows with experiments in mid-phase
      elev_CO2_exp |>                                                # Elevated CO2 Experiments data set
        filter(exp_time >= 3 & exp_time <= 6) |>                     # Take rows where experiment runs between 3 and 6 years
        mutate(exp_phase_name = "mid-phase (3-6 years since start)", # Add column with experiment phase name
               exp_phase_num = 2)) |>                                # Add column with experiment phase number
    bind_rows(                                                       # Combine with rows with experiments in late
      elev_CO2_exp |>                                                # Elevated CO2 Experiments data set
        filter(exp_time > 6) |>                                      # Take rows where experiment runs longer than 6 years
        mutate(exp_phase_name = "late phase (>6 years since start)", # Add column with experiment phase name
               exp_phase_num = 3)                                    # Add column with experiment phase number
    )

# Calculating the log-response ration per time phase
phase_LRR <- elev_CO2_exp |>                                 # Elevated CO2 Experiments data set
  select(starts_with("exp_phase"), ends_with("CO2_mean")) |> # Take columns from interest
  mutate(LRR = log(inc_CO2_mean / amb_CO2_mean)) |>          # Calculate Log-Response Ration for each measurement
  group_by(exp_phase_name) |>                                # Group data by experiment
  summarise(LRR_mean = mean(LRR),                            # Calculate mean values per experiment
            exp_phase_num = mean(exp_phase_num)) |>          # Experiment phase number
  arrange(exp_phase_num) |>                                  # Arrange by experiment phase number
  select(-exp_phase_num)                                     # Drop experiment phase number column (no longer used)

# Print table with results
phase_LRR |>
  knitr::kable(digits = c(NA, 4),
               col.names = c("Experiment Phase",
                             "Log-response ratio"),
               format = "html") |>
  kableExtra::kable_classic(full_width = F)
```
